---
title: "Exploring Tweedie Fit"
author: "Peter Novák"
date: "23/6/2022"
output:
  html_document:
    df_print: paged
  pdf_document:
    fig_height: 3.7
    fig_width: 10
geometry: a4paper, margin = 1.7cm
fontsize: 11pt
---

# Data Exploration - Bingo Aloha

## Get packages:

```{r, warnings = F, message = F, echo = F}
# pymc3
options(gsubfn.engine = "R")
library(tweedie)
library(ggplot2)
library(gridExtra)
library(fitdistrplus)
library(plotly)
library(sqldf)
```

```{r, warnings = F, message = F, echo = F}
dtweedie = function (x, mu, phi, power = NULL) {
    y = x
    xi = NULL
    xi.notation <- TRUE
    if (is.null(power)) {
        power <- xi
    }
    else {
        xi.notation <- FALSE
    }
    if (is.null(xi)) {
        xi.notation <- FALSE
        xi <- power
    }
    index.par <- ifelse(xi.notation, "xi", "p")
    index.par.long <- ifelse(xi.notation, "xi", "power")
    mu <- array(dim = length(y), mu)
    if (length(phi) > 1) {
        if (length(phi) != length(y)) 
            stop("phi must be scalar, or the same length as y.\n")
    }
    else {
        phi <- array(dim = length(y), phi)
    }
    density <- y
    if (power == 3) {
        density <- statmod::dinvgauss(x = y, mean = mu, dispersion = phi)
        return(density)
    }
    if (power == 2) {
        density <- dgamma(rate = 1/(phi * mu), shape = 1/phi, 
            x = y)
        return(density)
    }
    if (power == 0) {
        density <- dnorm(mean = mu, sd = sqrt(phi), x = y)
        return(density)
    }
    if ((power == 1) & (all(phi == 1))) {
        density <- dpois(x = y/phi, lambda = mu/phi)
        return(density)
    }
    id.type0 <- array(dim = length(y))
    id.series <- id.type0
    id.interp <- id.type0
    id.type0 <- (y == 0)
    if (any(id.type0)) {
        if (power > 2) {
            density[id.type0] <- 0
        }
        else {
            lambda <- mu[id.type0]^(2 - power)/(phi[id.type0] * 
                (2 - power))
            density[id.type0] <- exp(-lambda)
        }
    }
    xi <- array(dim = length(y))
    xi[id.type0] <- 0
    xi[!id.type0] <- phi[!id.type0] * y[!id.type0]^(power - 2)
    xix <- xi/(1 + xi)
    if ((power > 1) && (power <= 1.1)) {
        id.series <- (!id.type0)
        if (any(id.series)) {
            density[id.series] <- dtweedie.series(y = y[id.series], 
                mu = mu[id.series], phi = phi[id.series], power = power)
        }
        return(density = density)
    }
    if (power == 1) {
        id.series <- rep(TRUE, length(id.series))
        id.interp <- rep(FALSE, length(id.series))
    }
    if ((power > 1.1) && (power <= 1.2)) {
        id.interp <- ((xix > 0) & (xix < 0.1))
        id.series <- (!(id.interp | id.type0))
        if (any(id.interp)) {
            grid <- stored.grids(power)
            p.lo <- 1.1
            p.hi <- 1.2
            xix.lo <- 0
            xix.hi <- 0.1
            np <- 15
            nx <- 25
        }
    }
    if ((power > 1.2) && (power <= 1.3)) {
        id.interp <- ((xix > 0) & (xix < 0.3))
        id.series <- (!(id.interp | id.type0))
        if (any(id.interp)) {
            grid <- stored.grids(power)
            p.lo <- 1.2
            p.hi <- 1.3
            xix.lo <- 0
            xix.hi <- 0.3
            np <- 15
            nx <- 25
        }
    }
    if ((power > 1.3) && (power <= 1.4)) {
        id.interp <- ((xix > 0) & (xix < 0.5))
        id.series <- (!(id.interp | id.type0))
        if (any(id.interp)) {
            grid <- stored.grids(power)
            p.lo <- 1.3
            p.hi <- 1.4
            xix.lo <- 0
            xix.hi <- 0.5
            np <- 15
            nx <- 25
        }
    }
    if ((power > 1.4) && (power <= 1.5)) {
        id.interp <- ((xix > 0) & (xix < 0.8))
        id.series <- (!(id.interp | id.type0))
        if (any(id.interp)) {
            grid <- stored.grids(power)
            p.lo <- 1.4
            p.hi <- 1.5
            xix.lo <- 0
            xix.hi <- 0.8
            np <- 15
            nx <- 25
        }
    }
    if ((power > 1.5) && (power < 2)) {
        id.interp <- ((xix > 0) & (xix < 0.9))
        id.series <- (!(id.interp | id.type0))
        if (any(id.interp)) {
            grid <- stored.grids(power)
            p.lo <- 1.5
            p.hi <- 2
            xix.lo <- 0
            xix.hi <- 0.9
            np <- 15
            nx <- 25
        }
    }
    if ((power > 2) && (power < 3)) {
        id.interp <- ((xix > 0) & (xix < 0.9))
        id.series <- (!(id.interp | id.type0))
        if (any(id.interp)) {
            grid <- stored.grids(power)
            p.lo <- 2
            p.hi <- 3
            xix.lo <- 0
            xix.hi <- 0.9
            np <- 15
            nx <- 25
        }
    }
    if ((power >= 3) && (power < 4)) {
        id.interp <- ((xix > 0) & (xix < 0.9))
        id.series <- (!(id.interp | id.type0))
        if (any(id.interp)) {
            grid <- stored.grids(power)
            p.lo <- 3
            p.hi <- 4
            xix.lo <- 0
            xix.hi <- 0.9
            np <- 15
            nx <- 25
        }
    }
    if ((power >= 4) && (power < 5)) {
        id.interp <- ((xix > 0) & (xix < 0.9))
        id.series <- (!(id.interp | id.type0))
        if (any(id.interp)) {
            grid <- stored.grids(power)
            p.lo <- 4
            p.hi <- 5
            xix.lo <- 0
            xix.hi <- 0.9
            np <- 15
            nx <- 25
        }
    }
    if ((power >= 5) && (power < 7)) {
        id.interp <- ((xix > 0) & (xix < 0.5))
        id.series <- (!(id.interp | id.type0))
        if (any(id.interp)) {
            grid <- stored.grids(power)
            p.lo <- 5
            p.hi <- 7
            xix.lo <- 0
            xix.hi <- 0.5
            np <- 15
            nx <- 25
        }
    }
    if ((power >= 7) && (power <= 10)) {
        id.interp <- ((xix > 0) & (xix < 0.3))
        id.series <- (!(id.interp | id.type0))
        if (any(id.interp)) {
            grid <- stored.grids(power)
            p.lo <- 7
            p.hi <- 10
            xix.lo <- 0
            xix.hi <- 0.3
            np <- 15
            nx <- 25
        }
    }
    if (power > 10) {
        id.series <- (y != 0)
        id.interp <- (!(id.series | id.type0))
    }
    if (any(id.series)) {
        density[id.series] <- dtweedie.series(y = y[id.series], 
            mu = mu[id.series], phi = phi[id.series], power = power)
    }
    if (any(id.interp)) {
        dim(grid) <- c(nx + 1, np + 1)
        rho <- dtweedie.interp(grid, np = np, nx = nx, xix.lo = xix.lo, 
            xix.hi = xix.hi, p.lo = p.lo, p.hi = p.hi, power = power, 
            xix = xix[id.interp])
        dev <- tweedie.dev(power = power, mu = mu[id.interp], 
            y = y[id.interp])
        front <- rho/(y[id.interp] * sqrt(2 * pi * xi[id.interp]))
        density[id.interp] <- front * exp(-1/(2 * phi[id.interp]) * 
            dev)
    }
    density
}

t_col = function(color, percent = 60, name = NULL) {
  rgb.val = col2rgb(color)
  t.col = rgb(rgb.val[1], rgb.val[2], rgb.val[3],
              max = 255,
              alpha = (100 - percent) * 255 / 100,
              names = name)
  invisible(t.col)
}

get_dist = function(
  input_list,
  distributions,
  distribution_name,
  colors = NULL,
  plot_output = TRUE,
  verbose = FALSE,
  depricate = TRUE
) {
  input_len = length(input_list)
  if ((input_len > 10000) && (depricate == TRUE)) {
    input_list = input_list[sample(1:input_len, 10000)]
  }
  
  par(mar = c(4, 3, 3, 3), mgp = c(2, 0.5,0), mfrow = c(2,2))
  dist_obj = NULL
  n = 1
  for (distr in distribution_name) {
    temp_fit = NULL
    if (distr == "tweedie") {
      temp_fit = fitdist(data   = input_list,
                         method = "mle",
                         distr  = "tweedie",
                         start  = list(power = 3, mu = 20, phi = 0.2),
                         lower  = c(1, 0, 0))
    }
    else {
      temp_fit = fitdist(data   = input_list,
                         method = "mle",
                         distr  = distr)
    }
    dist_obj[[n]] = temp_fit
    n = n + 1
  }
  if (plot_output) {
    denscomp(dist_obj,
             legendtext = distribution_name,
             fitcol = colors)
    qqcomp(dist_obj,
           legendtext = distribution_name,
           fitcol = unlist(lapply(colors, function(x) t_col(x))),
           fitpch = 16, cex = 0.8)
    cdfcomp(dist_obj,
            xlogscale = T, ylogscale = F,
            legendtext = distribution_name,
            fitcol = colors)
    ppcomp(dist_obj,
           legendtext = distribution_name,
           fitcol = unlist(lapply(colors, function(x) t_col(x))),
           fitpch = 16, cex = 0.6)
  }
  
  gof_list = lapply(dist_obj, function(x) { gofstat(x)$aic })
#  return(gof_list)
#}

  if (verbose) {
    print("")
    for (i in 1:length(distribution_name)) {
      print(paste0("Goodness of Fit Values for ", distribution_name[i], ":"))
      print(gofstat(dist_obj[[i]]))
      print("")
    }
  }
  return(list(
    param_est  = dist_obj[[which.min(gof_list)]]$estimate,
    model_used = distribution_name[[which.min(gof_list)]],
    model_fun = distributions[[which.min(gof_list)]]
    ))
}
```

## Load and clean data:

```{r}
df_all = read.csv("./data/test_data_ba_1.csv")
df_all$user_id    = as.character(df_all$user_id)
df_all$test_group = as.character(df_all$test_group)
df_all$is_payer   = ifelse(df_all$total_spend == 0, 0, 1)
head(df_all)

df_payers = df_all[df_all$is_payer == 1, ]
head(df_payers)
```

## Clear Problem of Distribution of Player Spend:

One reason for this problem is the presence of duplicate values in the distribution, even worse in winsorised spend as it represents a higher proportion of total player spend:

```{r, echo = F}
# Total Spend:
sqldf("
      WITH t1 AS (
        SELECT COUNT(*)         count_of
             , SUM(total_spend) sum_of
             , total_spend
             , test_group
        FROM df_payers
        GROUP BY total_spend
               , test_group
      )
      SELECT CASE
                  WHEN count_of > 1 THEN 1
                  ELSE 0
             END           dup_spend
           , test_group
           , SUM(count_of) count_of
           , SUM(sum_of)   sum_of
      FROM t1
      GROUP BY CASE WHEN count_of > 1 THEN 1 ELSE 0 END
             , test_group
      ORDER BY test_group
             , dup_spend
      ")

# Win Spend:
sqldf("
      WITH t1 AS (
        SELECT COUNT(*)              count_of
             , SUM(total_wins_spend) win_sum_of
             , total_wins_spend
             , test_group
        FROM df_payers
        GROUP BY total_wins_spend
               , test_group
      )
      SELECT CASE
                  WHEN count_of > 1 THEN 1
                  ELSE 0
             END             dup_spend
           , test_group
           , SUM(count_of)   count_of
           , SUM(win_sum_of) win_sum_of
      FROM t1
      GROUP BY CASE WHEN count_of > 1 THEN 1 ELSE 0 END
             , test_group
      ORDER BY test_group
             , dup_spend
      ")
```

To visualise and compare to a hypothesised continuous distribution which will always assume the probability of an exact value is 0 thus the probability the same value happens twice in a finite sampled data set is 0 as well.

```{r, echo = F}
par(mar = c(4, 3, 3, 3)-0.2, mgp = c(2, 0.5,0), mfrow = c(1,1))
win_spend = df_payers$total_wins_spend


hist(log(win_spend), breaks = length(win_spend)/20)


par(mar = c(4, 3, 3, 3)-0.2, mgp = c(2, 0.5,0), mfrow = c(2,1))
spend_lnorm = rlnorm(length(win_spend),
                     get_dist(df_payers$total_spend, c(rnorm), c("lnorm"), F, F)$param_est)
value_frequency_obs   = as.numeric(as.character(as.data.frame(table(win_spend))$Freq))
value_frequency_lnorm = as.numeric(as.character(as.data.frame(table(spend_lnorm))$Freq))

plot(value_frequency_obs,
     type = "l",
     main = "Observed Value Frequency Distribution",
     ylab = "P(Spend=X)")
plot(value_frequency_lnorm,
     type = "l",
     main = "Expected Value Frequency Distribution",
     ylab = "P(Spend=X)")
```

So how well do most distributions do? The distribution is closest to tweedie out of them:

```{r}
distributions = c(rlnorm, rexp, rtweedie)
distribution_name = c("lnorm", "exp", "tweedie")
dist_obj = get_dist(df_payers$total_wins_spend, distributions, distribution_name,
                    c("plum", "dodgerblue", "darkgreen"), T, F)
dist_obj[1:2]
```

And all criterions outperform log-normal distribution with opt. parameters. p-value of ks-test 0.07

```{r, eval = F, echo = F}
temp_fit = fitdist(data = df_payers$total_wins_spend,
                   method = "mle",
                   distr  = "tweedie",
                   start  = list(power = 3, mu = 20, phi = 0.2),
                   lower  = c(1, 0, 0))
gofstat(temp_fit)
```

Conclusion:

The tweedie distribution is a better fit for our data, however the presence of duplicate spend values will make any unimodal distribution a likely poor fit. However it would have to be a highly irregular distribution because of the nature of the data (explained logic below):

1) User makes a spend by buying one of x packages with a certain value.

  a) This value is likely to be a "clean" value in the local currency (e.g. 0.99\$, 1.99\$, 4.99\$, 9.99\$, 49.99\$, 99.99\$)
  
  b) Local currency is translated to USD but those spend combination are likely to be duplicitous as well.
  
  c) a & b repeats when the same player makes a subsequent transaction which adds up to total spend per user.
  
2) All spends per user are summed up to make a per user total spent.

3) Daily spends are winsorized -> daily spend above certain threshold is set to threshold -> likely to create duplicitous total spend per user if user only spends on one day and spends above limit.

This process leads to dupl. spend values.

# So what now?

**One caveat, is assuming the data follows a exponential distribution even wrong?**

"One worry might be that selecting a model after considering the data is HARKing (hypothesizing after the results are known; Kerr 1998). In that famous article, Kerr discusses why HARKing may be inadvisable. In particular, HARKing can transform Type I errors (false alarms) into confirmed hypothesis ergo fact. I.e. the non-linear trend in the population data might be a random fluke, and the better fit by the non-linear distribution might be a random fluke.

Bayesian analysis is always conditional on the assumed model space. Often the assumed model space is merely a convenient default. The default is convenient because it is familiar to both the analyst and the audience of the analysis, but the default need not be a theoretical commitment. There are also different goals for data analysis: Describing the one set of data in hand, and generalizing to the population from which the data were sampled. Various methods for penalizing overfitting of noise are aimed at finding a statistical compromise between describing the data in hand and generalizing to other data."

**Further:**

"Indeed, according to Kleijn, v.d Vaart (2012), in the misspecified case, the posterior distribution:

* converges as $n \to \inf$ to a Dirac distribution cantered at a $\hat{\theta}_{ML}$

* does not have the correct variance (unless two values just happen to be same) in order to ensure that credible intervals of the posterior match confidence intervals for $\theta$."

**See for papers about updating prior distributions alongside prior parameters when presented with new data:**

* https://rss.onlinelibrary.wiley.com/doi/full/10.1111/rssb.12158

* https://xianblog.wordpress.com/2013/07/15/a-general-framework-for-updating-belief-functions/

### Next steps:

Investigate best fit distribution for each client data set.

Create a "population" of spenders by analysing every dataset we have so far about player spend.

# Lets try it for each firm:

```{r}
load_df = function(file_path) {
  df_all = read.csv(file_path)
  df_all$user_id    = as.character(df_all$user_id)
  df_all$test_group = as.character(df_all$test_group)
  df_all$is_payer   = ifelse(df_all$total_spend == 0, 0, 1)

  df_payers = df_all[df_all$is_payer == 1, ]
  return(df_payers)
}


visualise_df = function(df, name) {
  name          = gsub(" ", "\n", name)
  n_breaks      = max(min(nrow(df)/200, 40), 20)
  temp_dens     = density(df$total_wins_spend, adjust = 2)
  temp_dens_log = density(log(df$total_wins_spend), adjust = 2)
  
  hist(df$total_wins_spend,
    col    = "floralwhite",
    breaks = n_breaks,
    prob   = TRUE,
    xlab   = "Payer Winz. Spend",
    main   = "Hist.: Payer Winz Spend")
  lines(temp_dens,
    col = "firebrick4")
  legend("topright", legend = name, bty = "n")

  hist(log(df$total_wins_spend),
    col    = "floralwhite",
    breaks = n_breaks,
    prob   = TRUE,
    xlab   = "Payer log(Winz. Spend)",
    main   = "Hist.: Payer ln(Winz Spend)")
  lines(temp_dens_log,
    col = "firebrick4")
  legend("topright", legend = name, bty = "n")
}

descdist_to_list = function(descdist_obj) {
  l_out = c(descdist_obj$min,
            descdist_obj$max,
            descdist_obj$median,
            descdist_obj$mean,
            descdist_obj$sd,
            descdist_obj$skewness,
            descdist_obj$kurtosis)
  
  return(l_out)
}
```

```{r}
df_bingo_aloha   = load_df("./data/data_bingo_aloha.csv")
df_homw          = load_df("./data/data_homw.csv")
df_idle_mafia    = load_df("./data/data_idle_mafia.csv")
df_spongebob     = load_df("./data/data_spongebob.csv")
df_terra_genesis = load_df("./data/data_terra_genesis.csv")
df_ultimex       = load_df("./data/data_ultimex.csv")
```

```{r}
par(mar = c(4, 3, 3, 3), mgp = c(2, 0.5,0))
graphics::layout(mat = matrix(1:6, nrow = 2, ncol = 3, byrow = F),
       heights = rep(1, 6), widths = rep(1, 6))

visualise_df(df_bingo_aloha, "Bingo Aloha")
visualise_df(df_homw, "HOMW")
visualise_df(df_idle_mafia, "Idle Mafia")
visualise_df(df_spongebob, "Spongebob")
visualise_df(df_terra_genesis, "Terra Genesis")
visualise_df(df_ultimex, "Ultimate X-Poker")
```

```{r}
n_boots = 500

dist_bingo_aloha   = descdist(df_bingo_aloha$total_wins_spend, boot = n_boots)
dist_homw          = descdist(df_homw$total_wins_spend, boot = n_boots)
dist_idle_mafia    = descdist(df_idle_mafia$total_wins_spend, boot = n_boots)
dist_spongebob     = descdist(df_spongebob$total_wins_spend, boot = n_boots)
dist_terra_genesis = descdist(df_terra_genesis$total_wins_spend, boot = n_boots)
dist_ultimex       = descdist(df_ultimex$total_wins_spend, boot = n_boots)

mat_out = matrix(data = c(descdist_to_list(dist_bingo_aloha),
                          descdist_to_list(dist_homw),
                          descdist_to_list(dist_idle_mafia),
                          descdist_to_list(dist_spongebob),
                          descdist_to_list(dist_terra_genesis),
                          descdist_to_list(dist_ultimex)),
                 nrow = 6, ncol = 7, byrow = T,
                 dimnames = list(c("Bingo Aloha",
                                   "HOMW",
                                   "Idle Mafia",
                                   "Spongebob",
                                   "Terra Genesis",
                                   "Ultimate X-Poker"),
                                 c("Min",
                                   "Max",
                                   "Median",
                                   "Mean",
                                   "SD",
                                   "Skewness",
                                   "Kurtosis")))
as.data.frame(mat_out)
```

```{r}
distributions = c(rlnorm, rweibull, rexp)#, rtweedie)
distribution_name = c("lnorm", "weibull", "exp")#, "tweedie")
colours = c("plum", "darkgreen", "dodgerblue")#, "cyan")

fit_bingo_aloha   = get_dist(df_bingo_aloha$total_wins_spend,
                             distributions, distribution_name, colours, T, F, F)
fit_homw          = get_dist(df_homw$total_wins_spend/100,
                             distributions, distribution_name, colours, T, F, F)
fit_idle_mafia    = get_dist(df_idle_mafia$total_wins_spend/100,
                             distributions, distribution_name, colours, T, F, F)
fit_spongebob     = get_dist(df_spongebob$total_wins_spend,
                             distributions, distribution_name, colours, T, F, F)
fit_terra_genesis = get_dist(df_terra_genesis$total_wins_spend,
                             distributions, distribution_name, colours, T, F, F)
fit_ultimex       = get_dist(df_ultimex$total_wins_spend,
                             distributions, distribution_name, colours, T, F, F)

mat_out2 = matrix(data = c(fit_bingo_aloha$model_used,
                           fit_homw$model_used,
                           fit_idle_mafia$model_used,
                           fit_spongebob$model_used,
                           fit_terra_genesis$model_used,
                           fit_ultimex$model_used),
                  nrow = 6, ncol = 1, byrow = T,
                  dimnames = list(c("Bingo Aloha",
                                    "HOMW",
                                    "Idle Mafia",
                                    "Spongebob",
                                    "Terra Genesis",
                                    "Ultimate X-Poker"),
                                  c("Model Used")))
as.data.frame(mat_out2)
```

All get lnorm as best dist.
